---
title: "Project2"
author: "Lin Li"
output: rmarkdown::github_document
params:
  weekday: 1
---
# Introduction
For this project, I will use bike sharing data to do analysis. There are 15 variables :  
-instant(record index)  
-dteday(date)  
-season(1:winter, 2:spring, 3:summer, 4:fall)  
-yr(0: 2011, 1:2012)  
-mnth(month: 1 to 12)  
-hr(hour:0 to 23)  
-holiday(weather day is holiday or not)  
-weekday(day of week)  
-workingday(1 for neither weekend nor holiday,0 otherwise)  
-weathersit(1: Clear, Few clouds, Partly cloudy, Partly cloudy,2: Mist and Cloudy, Mist andBroken clouds, Mist and Few clouds, Mist,3: Light Snow, Light Rain and Thunderstorm and Scattered clouds, Light Rain and Scattered clouds,4: Heavy Rain and Ice Pallets and Thunderstorm and Mist, Snow and Fog )  
-temp(Normalized temperature in Celsius)  
-atemp( Normalized feeling temperature in Celsius)  
-hum(Normalized humidity),windspeed(Normalized wind speed)  
-cnt(count of total rental bikes including both casual and registered ) 

The purpose of this analysis is to create models for predicting cnt variable using others predictors mentioned above. Two models are created:classification tree model using leave one out cross validation, boosted tree model using cross validation.




```{r,warning=FALSE,include=FALSE}
# library used 
library(tidyverse)
library(caret)
library(rmarkdown)
library(corrplot)
library(knitr)
```


# Data
```{r}
# relative path to read data
#getwd()
day<-read.csv('day.csv')
day<-day%>%select(-c('casual','registered'))
hour<-read.csv('hour.csv')
hour<-hour%>%select(-c('casual','registered'))
# combine two data and get a new one
data<-semi_join(hour,day,by='dteday')
# filter Monday out
date<- filter(data,weekday==params$weekday)
set.seed(100)
# set train/test set
index<-createDataPartition(date$cnt,p=0.7,list=FALSE)
trainday<-date[index,]
testday<-date[-index,]

```


# Summarization

In this section, I summarize information for Monday, draw density plots for all variables, and plot a correlation plot for cnt with others predictors.  From density plots, I can conclude the distributions for variables. From correlation plot, I can conclude correlation relationships between each variable.

```{r,warning=FALSE}
# summary for monday
summary(date)
# density plot
date %>%keep(is.numeric) %>%pivot_longer(everything()) %>%ggplot(aes(x = value)) +facet_wrap(~ name, scales = "free")+
geom_density()
# all predictors' correlation with cnt variable
correlation <- cor(select(date,season,yr,mnth,holiday,workingday,weathersit,temp,atemp,hum,windspeed,cnt))
corrplot(correlation, type = "upper", tl.pos = "lt",use="pairwise.complete.obs")
corrplot(correlation, type = "lower", method = "number", add = TRUE, tl.pos = "n",use="pairwise.complete.obs")
```


# Modeling

Two models are created for predicting cnt of Monday: First is classification model using leave one out cross validation. Second is boosted tree model with cross validation. I read in data and split into training and test.Then centered and scaled data, and fitted models using methods mentioned.  
Through comparing RMSE for test dataset, choose the one with smaller RMSE as final model.

```{r,warning=FALSE}
# fit model 
classification.fit<-train(cnt~season+mnth+hr+holiday+workingday+weathersit+temp+atemp+hum+windspeed,trainday,method='rpart',preProcess = c('center','scale'),trControl=trainControl(method='LOOCV'))
classification.fit
boosted.fit<-train(cnt~season+yr+mnth+hr+holiday+workingday+weathersit+temp+atemp+hum+windspeed,trainday,method='gbm',preProcess = c('center','scale'),trControl=trainControl(method='repeatedcv'), verbose = FALSE)
boosted.fit
# predict
class<-predict(classification.fit,newdata = testday)
classRMSE<-sqrt(mean((class-testday$cnt)^2))
boost<-predict(boosted.fit,newdata = testday)
boostRMSE<-sqrt(mean((boost-testday$cnt)^2))
#RMSE for classification model and boost model
x<-c(classification=classRMSE,boost=boostRMSE)
knitr::kable(x,col.names =' RMSE')
```






